Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-18 main.tex
tiny changes

  We have the following derivations:
  $$
  \vlderivation{
    \vlin{\cD}{}{(A \vlor B) \vlan (C \vlor D)}{
      \vlin{\wlD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((A \vlor B) \vlan (C \vlor D))}{
	\vlin{\wlD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((A \vlor B) \vlan D))}{
	  \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor (B \vlan D))}{
	    \vlin{\wrD}{}{((A \vlor B) \vlan C) \vlor (B \vlan D)}{
	      \vlhy{(A \vlan C) \vlor (B \vlan D)}}}}}}}
  $$.
  $$
  \vlderivation{
    \vlin{\cD}{}{\exists x.(A \vlor B)}{
      \vlin{\wlD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.(A \vlor B))}{
	\vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.B)}{
	  \vlhy{(\exists x.A) \vlor (\exists x.B)}}}}}
  $$.
  $$
  \vlderivation{
    \vlin{\cD}{}{\forall x.(A \vlor B)}{
      \vlin{\wlD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(A \vlor B))}{
	\vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.B)}{
	  \vlhy{(\forall x.A) \vlor (\forall x.B)}}}}}
  $$.
  \juihsuan{done. If needed, we can introduce the notion of open deduction to
reduce the size of derivations...}\lutz{I was thinking about that, but (i) it is probably not worth the effort, as we wont have many derivation, and (ii) it is hard to define rectified derivations this way.}
\end{proof}

@@

For a derivation $\Deri$ in
$\set{\wlD,\wrD,\acD,\me,\mfaD,\mexD,\fequ}$, we can now construct the
\bfit{rectification} $\rectif\Deri$ by rectifying each line of $\Deri$, yielding
a derivation in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$.

For each instance $\vlinf{\rr}{}{P}{Q}$ of an inference rule in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can
define the \bfit{induced map} $\mapof{\rr}\colon\vgraphof
Q\to\vgraphof P$ which acts as the identity for
$\rr\in\set{\me,\fequ}$ and as the canonical injection for
$\rr\in\set{\wlD,\wrD}$. For $\rr=\acD$ it maps the vertices
corresponding to the two atoms in the premise to vertex of the
contracted atom in the conclusion, and for
$\rr\in\set{\rectif\mfaD,\rectif\mexD}$ it maps the two vertices
corresponding to the quantifiers in the premise to the one in the
conclusion (as acts as the identity on all other vertices). For a
derivation $\Deri$ in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can then define
the \bfit{induced map} $\mapof\Deri$ as the composition of the induced
maps of the rule instances in $\Deri$. \juihsuan{maybe mention at least that the induced
maps define graph homomorphisms. Do we need to talk about the contexts
$\Sconhole$ here (induced maps act clearly as the identity on contexts but we need them for the composition)?}

>>>>

  We have the following derivations:
  $$
  \vlderivation{
    \vlin{\cD}{}{(A \vlor B) \vlan (C \vlor D)}{
      \vlin{\fequ}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((A \vlor B) \vlan (C \vlor D))}{
      \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((B \vlor A) \vlan (D \vlor C))}{
	\vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor ((B \vlor A) \vlan D))}{
	  \vlin{\wrD}{}{((A \vlor B) \vlan (C \vlor D)) \vlor (B \vlan D))}{
	    \vlin{\wrD}{}{((A \vlor B) \vlan C) \vlor (B \vlan D)}{
	      \vlhy{(A \vlan C) \vlor (B \vlan D)}}}}}}}}
  $$.
  $$
  \vlderivation{
    \vlin{\cD}{}{\exists x.(A \vlor B)}{
      \vlin{\fequ}{}{\exists x.(A \vlor B)) \vlor (\exists x.(A \vlor B))}{
      \vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.(B \vlor A))}{
	\vlin{\wrD}{}{(\exists x.(A \vlor B)) \vlor (\exists x.B)}{
	  \vlhy{(\exists x.A) \vlor (\exists x.B)}}}}}}
  $$.
  $$
  \vlderivation{
    \vlin{\cD}{}{\forall x.(A \vlor B)}{
      \vlin{\fequ}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(A \vlor B))}{
      \vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.(B \vlor A))}{
	\vlin{\wrD}{}{(\forall x.(A \vlor B)) \vlor (\forall x.B)}{
	  \vlhy{(\forall x.A) \vlor (\forall x.B)}}}}}}
  $$.
  \juihsuan{done. If needed, we can introduce the notion of open deduction to
reduce the size of derivations...}\lutz{I was thinking about that, but (i) it is probably not worth the effort, as we wont have many derivation, and (ii) it is hard to define rectified derivations this way.}
\end{proof}

@@

For a derivation $\Deri$ in
$\set{\wrD,\acD,\me,\mfaD,\mexD,\fequ}$, we can now construct the
\bfit{rectification} $\rectif\Deri$ by rectifying each line of $\Deri$, yielding
a derivation in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$.

For each instance $\vlinf{\rr}{}{P}{Q}$ of an inference rule in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can
define the \bfit{induced map} $\mapof{\rr}\colon\vgraphof
Q\to\vgraphof P$ which acts as the identity for
$\rr\in\set{\me,\fequ}$ and as the canonical injection for
$\rr=\wrD$. For $\rr=\acD$ it maps the vertices
corresponding to the two atoms in the premise to vertex of the
contracted atom in the conclusion, and for
$\rr\in\set{\rectif\mfaD,\rectif\mexD}$ it maps the two vertices
corresponding to the quantifiers in the premise to the one in the
conclusion (as acts as the identity on all other vertices). For a
derivation $\Deri$ in
$\set{\wrD,\acD,\me,\rectif\mfaD,\rectif\mexD,\fequ}$ we can then define
the \bfit{induced map} $\mapof\Deri$ as the composition of the induced
maps of the rule instances in $\Deri$. \juihsuan{maybe mention at least that the induced
maps define graph homomorphisms. Do we need to talk about the contexts
$\Sconhole$ here (induced maps act clearly as the identity on contexts but we need them for the composition)?}
